---
title: "Ggplot2"
author: "Syed Faraz Ali"
date: "10/6/2018"
output:
  html_document:
    code_folding: show
    theme: readable
    toc: true
    toc_float:
      collapsed: no
      smooth_scroll: yes
      toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
library(ggplot2)
library(magrittr)
library(dplyr)
```

```{r}
str(mpg)
```


```{r}
qplot(x = displ, y = hwy, data = mpg)
```

Modifying aesthetics 

```{r}
qplot(x = displ, y = hwy, data = mpg, color = drv)
```


Adding a geom
```{r}
qplot(x = displ, y = hwy, data = mpg, geom = c("point", "smooth"))
```


Histograms

```{r}
qplot(x = hwy, data = mpg, fill = drv)
```

Facets 

```{r}
qplot(x = displ, y = hwy, data = mpg, facets = . ~ drv)
```

```{r}
qplot(x = hwy, data = mpg, facets = drv ~., binwidth = 2)
```

Reading maacs data 

```{r}
load("~/Downloads/maacs.Rda")
maacs %>% head()
maacs %>% str
```

Histogram of eNO 

```{r}
qplot(x = log(eno), data = maacs )
```


Histogram by groups 

```{r}
qplot(x = log(eno), data = maacs, fill = mopos)
```

Density smooth 

```{r}
qplot(x = log(eno), data = maacs, geom = "density")
qplot(x = log(eno), data = maacs, geom = "density", color = mopos)
```

Scatterplots 

```{r}
ggplot(data = maacs, mapping = aes(x = log(pm25), y = log(eno)))+
        geom_point()

ggplot(data = maacs, mapping = aes(x = log(pm25), y = log(eno), shape = mopos)) + 
        geom_point()

ggplot(data = maacs, mapping = aes(x = log(pm25), y = log(eno), color = mopos)) + 
        geom_point()+
        geom_smooth(method = "lm")
```


Using facets 

```{r}
gg <- ggplot(data = maacs, mapping = aes(x = log(pm25), y = log(eno))) +
        geom_point()+
        geom_smooth(method = "lm")
        
plot(gg)

qplot(x = log(pm25), y = log(eno), data = maacs, facets = .~mopos) + 
        geom_smooth(method = "lm")


# gg + facet_wrap(.~mopos)
```


Building up in layers 

```{r}
load("~/Downloads/maacs (1).Rda")
head(maacs[,6:8])
```

Plotting 

```{r}
load("~/Downloads/maacs (1).Rda")
g <- ggplot(data = maacs, mapping = aes(x = logpm25, y = NocturnalSympt))
summary(g)
plot(g)

p <- g + geom_point()
plot(p)
```

Adding more layers : Smooth and facets
```{r}
g + geom_point() + geom_smooth()
g + geom_point() + geom_smooth(method = "lm")
g + 
        geom_point()+
        facet_grid(. ~ bmicat)+
        geom_smooth(method = "lm")
```


Modifying aesthetics 

```{r}
g + geom_point(color = "steelblue", size = 4, alpha = 0.5)
g + geom_point(aes(color = bmicat), size = 4, alpha = 0.5)
```

Modifying labels 

```{r}
g +geom_point(aes(color = bmicat)) + 
        labs(title = "MAACS Cohort")+
        labs (x = expression("log "* PM[2.5]), y = "Nocturnal Symptoms")
```


Customizing the smooth 

```{r}
g + geom_point(aes(color = bmicat), size =2, alpha = 0.5)+
        geom_smooth(size = 4, linetype = 3, method = "lm", se = FALSE )
```

Changing the theme

```{r}
g + geom_point(aes(color = bmicat)) + 
        theme_bw(base_family = "Times")
```


Notes about Axis Limits 

With base plot
```{r}
testdat <- data.frame(x = 1:100, y = rnorm(100))
testdat[50,2] <- 100  # Outlier 
plot(testdat$x, testdat$y, type = "l", ylim = c(-3,3))
```

With ggplot 

```{r}
g <- ggplot(data = testdat, mapping = aes(x = x, y = y))
g +  geom_line()
```

Exploring the maacs data again 

```{r}
load("~/Downloads/maacs (1).Rda")
maacs %>% head

(cutpoints <- quantile(maacs$logno2_new, seq(0,1,length = 4), na.rm = TRUE))

maacs$no2dec <- cut(maacs$logno2_new,cutpoints) 

levels(maacs$no2dec)
```

Code for final plot 

```{r}
g <- ggplot(data = maacs, mapping = aes(x = logpm25, y = NocturnalSympt))

## Add layers

g + geom_point(alpha = 1/3)+
        facet_wrap(bmicat ~ no2dec, nrow = 2, ncol = 4)+
        geom_smooth(method = "lm", se = FALSE, col = "steelblue")+
        theme_bw(base_family = "Avenir", base_size = 10)+
        labs(x = expression("log " * PM[2.5]))+
        labs(y = "Nocturnal Symptoms")+
        labs(title = "MAACS Cohort")


```














